---
title: "Regression Final Project"
author: "Joe DeMaro"
date: "April 4, 2019"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(GGally)
data("mtcars")
```
## Executive Summary
You work for Motor Trend, a magazine about the automobile industry. 
Looking at a data set of a collection of cars, 
they are interested in exploring the relationship between 
a set of variables and miles per gallon (MPG) (outcome). 
They are particularly interested in the following two questions:

"Is an automatic or manual transmission better for MPG"

"Quantify the MPG difference between automatic and manual transmissions"

To investigate the first question, a comparison of the raw data is used to drive a linear regression comparsion between the two variables within the inquire. The inital analysis shows a 7.25 mpg difference between manual and automatic transmissson.  The p-value for the comparison is 0.0014 indicating the difference is significant.  The regression data indicates this comparison only account for ~35% of the differences.
The initial pass does not help determine if any other measurement affects mpg.  To determine if other factors might affect MPG, a regression model is executed with all variables considered. The information from this model (mod1) led to significant differences being seen between mpg and wgt, cyl, and disp.  These 3 variables were modeled with am to identify other variables affecting mpg.  wgt and cyl significantly affect mpg and the data show the 4 variables account for over 80% of the differences in mpg
Residual analysis plot show that the data are non problematic and linear regression modeling is appropriate.

# Appendix
## Exploratory Data Visualization
```{r mtcars}

head(mtcars)
str(mtcars)

data.subset <- subset(mtcars, select=c(mpg,am))
mt.auto <- mtcars[(mtcars$am==0),]
mt.manual <- mtcars[(mtcars$am==1),]

boxplot(mpg ~ am, data=mtcars, xlab="Transmission", ylab="MPG", main="MPG v Transmission Type")
summary(mtcars)
```
## Compare Auto vs. Manual Transmission
```{r}
aggregate(data.subset[,1:2], list(data.subset$am), mean)
mod0 <- lm(mpg ~ am, mtcars)
summary(mod0)
t.test(mt.auto$mpg, mt.manual$mpg)
diffTrans <- mean(mt.manual$mpg) - mean(mt.auto$mpg)
### Difference between manual and automatic transmission
diffTrans
```
## Effect of Multiple Variables
### Initial comparison driven by the questions of interest provide limited insight into the mtcars data. With additional data available and the linear regression results of mod0 indicating only 35% of the difference in mpg can be explained by am, a more complete analysis will be gleaned by looking at a broad comparison of all the variables and then determining a more filtered view based on the additional information.
```{r}
mod1 <- lm(mpg~cyl + disp + hp + drat + wt + qsec + vs + am + gear + carb, data=mtcars)
summary(mod1)$coef
summary(aov(mod1))
```
## Deeper Analysis
### The analysis of variance from the regression model (mod1) shows 3 significantly different factors. A new model (mod2) below will limit x variables to the 3 discovered variables plus am, part of the original question.  The results show that transmission type plus number of cylinders and weight  have a significant impact on MPG
```{r}
mod2 <- lm(mpg ~ factor(am) + cyl + disp + wt, mtcars)
summary(mod2)
summary(aov(mod2))
anova(mod0,mod2)
```
## Residual Analysis
```{r}
par(mfrow=c(2,2))
plot(mod2)
```
