---
title: "Statistical Inference - Simulation Project"
author: "Joe DeMaro"
date: "December 2018"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Set Parameters
```{r}
setwd("c:\\rprograms\\statinference")
n <- 1000
group <- 40
lambda <- 0.2
```
##Calculate Theoretical Mean and SD
```{r}
theoreticalmean <- 1 / lambda
theoreticalsd <- (1 / (lambda * lambda)) / group
```
##Run Simulation
```{r echo=TRUE}
# run the simulation
data  <- rexp(n * group, lambda )
matrixdata <- matrix(data, n, group)
matrixmean <- apply(matrixdata,1,mean)
```
##Tabulate Simulation Data
```{r echo=TRUE}
simmean <- mean(matrixmean)
simsd <- sd(matrixmean)
simvar <- var(matrixmean)

#compare to CLT
simse = simsd / sqrt(group)
#calculate confidence
low <- simmean - 1.96 * simse
high <- simmean + 1.96 * simse

library(knitr)

theorecticalData <- c(theoreticalmean,theoreticalsd)

simulationdata <- c(simmean, simsd)

df = data.frame(theorecticalData,simulationdata)
colnames(df) <- c("Theorectical", "Simulated")
row.names(df) <- c("Mean", "SD")

kable(df, format="markdown", caption="Comparison between Theorecital and Simulated Data")
```
## Distribution of Simulated Data around the mean (95% CI)
```{r}
ciData <- c(low, high)
ciDF <- data.frame(ciData)
row.names(ciDF) <- c("Low", "High")
kable(ciDF, format="markdown", caption="CI for Simulated Data", title="Simulated Data CI")
```

## Histogram of Means from Simulation
```{r}
hist(matrixmean)
```

##Analysis

### The null hypothesis Ho is the mean = 5; making the alternate hypothesis Ha the mean != 5. With a 95% confidence level, the simulated data is generated to test.  The histogram shows the data are normally distrubtedd and the data show the simulated mean falls with in the 95% level.  Therefore you can not reject the Ho.