---
title: "PML Final Project"
author: "Joe DeMaro"
date: "9/11/2019"
output: html_document
---

# Abstract
Data from IoT devices for fitness collect an enormous amount of data about an individuals personal activities.  For this analysis, data from 6 individuals performing barbell lifts in a variety of correct and incorrect fashions was collected. 

The task for the analyst is to predict the manner in which a user did the exercise.  The given data consists or a training and a test group.  The data from both groups is cleaned and analyzed in a variety of fashions to determine the best model for prediction. The provided test set is held until the best model is created.  The training set is divided into a training and testing set to help determine the best model.



```{r echo=FALSE, warning=FALSE}
library(readr)
library(AppliedPredictiveModeling)
library(caret)
library(gbm)
library(corrplot)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
```

## Download Data
```{r dataload, warning=FALSE}
## load training data
training = read_csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"))
dim(training)

## load test data
testing <- read_csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"))
dim(testing)
```
## clean training data

```{r TrainingData}
indColToRemove <- which(colSums(is.na(training) | training=="")>0.9*dim(training)[1]) 
training_clean <- training[,-indColToRemove]
training_clean <- training_clean[,-c(1:7)]
dim(training_clean)



```

## Clean Model Verification Data
```{r FinalData}
indColToRemove <- which(colSums(is.na(testing) | testing=="")>0.9*dim(testing)[1]) 
final_clean <- testing[,-indColToRemove]
final_clean <- final_clean[,-c(1:7)]
dim(final_clean)
```

```{r TrainingTestSets}

set.seed(12345)
## Get training data
inTrain <- createDataPartition(training_clean$classe, p = 0.7, list = FALSE)
trainData <- training_clean[inTrain, ]
dim(trainData)
#Get Test Set
testData <- training_clean[-inTrain, ]
dim(testData)
```

## Correlation Analysis
```{r Correlation}
corr_data <- cor(trainData[, -53])
corrplot(corr_data, method="circle")

## find highly correlated columns
highCor <- findCorrelation(corr_data, cutoff = 0.80)
names(trainData)[highCor]

decisionTreeMod1 <- rpart(classe ~ ., data=trainData, method="class")
fancyRpartPlot(decisionTreeMod1)

predictTreeMod1 <- predict(decisionTreeMod1, testData, type = "class")

ConMat <- confusionMatrix(factor(predictTreeMod1), factor(testData$classe))
ConMat

#plot accuracy data
```
## Plot Accuracy for Decision Tree
```{r Plot1}
plot(ConMat$table, col=ConMat$byClass, 
      main = paste("Decision Tree Accuracy = ", round(ConMat$overall['Accuracy'],4)))
```
## Random Forest Analysis
```{r RFAnalysis}
controlRF <- trainControl(method="cv",number=3, verboseIter = FALSE)
modRF1 <- train(classe ~ ., data=trainData, method="rf", trControl=controlRF)
modRF1$finalModel
```

## Validate RF model
```{r RFValidate}
predictRF1 <- predict(modRF1, newdata=testData)
ConRF <- confusionMatrix(factor(predictRF1), factor(testData$classe))
ConRF

plot(modRF1)
```
## Plot Accuracy of Random Forest Model

```{r RFAccuracyPlot}

plot(ConRF$table, col=ConRF$byClass, main=paste("RF Confusion Matrix: Accuracy =", round(ConRF$overall['Accuracy'], 4)))
```
## GBM
```{r GBM}
set.seed(12345)
cGBM <- trainControl(method="repeatedcv", number=5,repeats=1)
modGBM <- train(classe ~ ., data=trainData, method="gbm", trControl=cGBM, verbose=FALSE)
modGBM$finalModel

print(modGBM)
```
# Validate GBM model
```{r GBMValidate}
predictGBM <- predict(modGBM, newdata = testData)

CMgbm <- confusionMatrix(factor(predictGBM), factor(testData$classe))
CMgbm
```

## Run test data provided in beginning from source.
```{r ValidateTestData}
Results <- predict(modRF1, newdata = final_clean)
Results
```
